---
title: "Assignment2"
author: "Luyi Huang"
date: "2021/2/25"
output:
  pdf_document:
    latex_engine: xelatex
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(GGally)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(car)
library(knitr)
library(devtools)
```

## Exercise 1 Data Desription
```{r}
library(bayesm)
data(margarine)
choiceprice=margarine$choicePrice
demos=margarine$demos
all=merge(choiceprice,demos,by="hhid")
```
```{r}
avg=apply(as.matrix(choiceprice[,3:12]),2, mean)
print(avg)
```
```{r}
std<-function(x) sd(x)/sqrt(length(x))
stdc=apply(as.matrix(choiceprice[,3:12]),2, std)
print(stdc)
```
```{r}
#Average and dispersion in product characteristics
library(tidyr)
library(dplyr)
#Compute all summary statistics including both chosen products and non-chosen products
means <- t(all %>% summarise_at(3:12,mean))
mins <- t(all %>% summarise_at(3:12,min))
maxs <- t(all %>% summarise_at(3:12,max))
sds <- t(all %>% summarise_at(3:12,sd))
vars <- t(all %>% summarise_at(3:12,var))
des1=cbind(means,mins,maxs,sds,vars)
label1=c("mean","min","max","sd","var")
colnames(des1) <- label1
des1=round(des1,digits=3)
des1
```
Market Share :
```{r}
ms<-as.data.frame(choiceprice %>% count(choice))
ms$n<-ms$n/nrow(all)
ms
```
Market Share by Characteristics:
```{r}
#finding the price 
for (i in 1:nrow(all)){
  all$choiceprice[i]=all[i,all$choice[i]+2]
  all$choices[i]=colnames(all)[all$choice[i]+2]
} 
```

```{r}
means <- as.data.frame((t(all %>% summarise_at(3:12,mean))))
means <- cbind(choices = rownames(means), means)
rownames(means) <- 1:nrow(means)
means
all=left_join(all, unique(means), by = c("choices"="choices"))
```
```{r}
all$indict<-all$choiceprice>all$V1
sum(all$indict)
```
```{r}
all%>%
  filter(all$indict==TRUE) %>%
  count(choice)%>%
  mutate(n=n/sum(all$indict))
```
```{r}
all%>%
  filter(all$indict==TRUE) %>%
  count(choice)%>%
  mutate(n=n/(nrow(all)-sum(all$indict)))
```
consider mapping on choices and attributes 
```{r}
des3= all %>%
  group_by(choice) %>%
  summarize(
    famsize1_2=sum(Fs3_4 == 0 & Fs5.==0),
    famsize3_4=sum(Fs3_4 == 1 & Fs5.==0),
    famsize5.=sum(Fs3_4 == 0 & Fs5.==1),
    college=sum(college==1),
    whtcollar=sum(whtcollar==1),
    retired=sum(retired==1)
  )
notdes3= all %>%
  group_by(choice) %>%
  summarize(
    notcollege=sum(college==0),
    notwhtcollar=sum(whtcollar==0),
    notretired=sum(retired==0)
  )
des3=merge(des3,notdes3)
des3
```
Larger familisize, more retired people will like choice1. Overall, choice1 is much more popular among all the attributes of customers.

## Exercise 2 First Model
consider using conditional logit model since the price value varies of the same alternate accross people.
Likelihood and Optimization: 
```{r}
library(mlogit)
library(stargazer)
library(texreg)
library(survival)
library(nnet)
price=margarine$choicePrice
colnames(price)[3:12]=paste0("price",1:10)
price<-price[-c(1)]
clogit0=mlogit.data(price,varying=2:11,shape="wide",sep="",choice="choice")
clogit1=mlogit(choice~price,data=clogit0)
summary(clogit1)
```
```{r}
ni=nrow(price)
nj=ncol(price[,2:11])
Y=class.ind(all$choice)
#consider likelihood, we only have one variable price and thus beta is a scalor in the function.
clogit_ll<-function(beta){
  
  #Create the constant as instructed
  intercept=cbind(0,matrix(rep(beta[1:nj-1],each=ni),ni,nj-1))
  
  #Use the lecture definition of conditional logit to compute the likelihood
  XB=price[,2:11]*beta[nj]
  XB=intercept+XB
  eXB=exp(XB)
  teXB=rowSums(eXB)
  prob=eXB/teXB
  
  #Compute the neg log likelihood for each choice using the choice matrix
  llik=sum(Y*log(prob))
  return(-llik)
}
```
```{r}
set.seed(0)
clogit <- optim(runif(10,-0.1,0.1),clogit_ll,method="BFGS")
```
We can see that the fixed effects of choices varies and increasing prices will reduce the probabiliy of purchasing.
```{r}
clogit$par
```
Interpretation:
Overall, the price has a negative impact on the probability of getting chosen by customers. 
## Exercise 3 Second Model
The second model should be the multinomial logit model since income varies across different household id.
```{r}
colnames(all)[3:12]=paste0("price",1:10)
mlogit0=mlogit.data(all,varying=3:12,shape="wide",sep="",choice="choice")
mlogit1=mlogit(choice~0 | Income,data=mlogit0)
summary(mlogit1)
```




```{r}
mlogit_ll<-function(beta){
  
  #Create the constant as instructed
  intercept1=cbind(0,matrix(rep(beta[1:9],each=ni),ni,nj-1))
  intercept2=cbind(0,matrix(rep(beta[10:18],each=ni),ni,nj-1))
  #Use the lecture definition of conditional logit to compute the likelihood
  XB=intercept1+intercept2*cbind(all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13])
  eXB=exp(XB)
  teXB=rowSums(eXB)
  prob=eXB/teXB
  
  #Compute the neg log likelihood for each choice using the choice matrix
  ll=-sum(Y*log(prob))
  return(ll)
}
```
```{r}
set.seed(0)
model2 <- optim(runif(18,-0.1,0.1),mlogit_ll,method="BFGS")
model2$par[10:18]
```
Income will have a negative impact of getting hired for choice2, choice7, while other choices have the attributes that if the income is higher, the probability of choosing will get higher.
## Exercise 4 Marginal Effects of First and Second Model
```{r}
beta=matrix(0,nrow=1,ncol=11)
beta[1,1]=0
beta[1,2:11]=clogit$par
delta<-diag(1,10,10)
```

```{r}
mf=matrix(0,nrow=10,ncol=10)
for (i in 1:nrow(all)){
  intercept=beta[1:10]
  
  #Use the lecture definition of conditional logit to compute the likelihood
  xibi=as.matrix(exp(price[i,2:11]*beta[11]+intercept))
  xib=rowSums(xibi)
  pij=xibi/xib
 dpij<-matrix(NA,10,10)
  delta<-diag(1,10,10)
 for (k in 1:10){
    delta[,k]=delta[,k]-pij
 }
 for (m in 1:10){
    dpij[m,]=pij*delta[m,]*beta[11]
 }
 mf=mf+dpij
}
```
```{r}
mf/4470
```


```{r}
z<-with(clogit0,data.frame(price=tapply(price,idx(clogit0,2),mean)))
effects(clogit1,covariate="price",data=z)
```
We can see that diagonal elements are almost negative, meaning that price will reduce the chance of getting chosen. The off-diagonal elements are almost positive, meaning that these products are subsititutes.
```{r}
beta=matrix(0,nrow=1,ncol=20)
beta[1,1]=0
beta[1,2:10]=model2$par[1:9]
beta[1,11]=0
beta[1,12:20]=model2$par[10:18]
```


Marginal Effects of MNL:
```{r}
intercept1=beta[1:10]
intercept2=beta[11:20]
mf=matrix(0,nrow=1,ncol=10)
for (i in 1:nrow(all)){
  xibi=exp(intercept1+intercept2*cbind(all[i,13],all[i,13],all[i,13],all[i,13],all[i,13],all[i,13],all[i,13],all[i,13],all[i,13],all[i,13]))
  xib=rowSums(xibi)
  pij=xibi/xib
  betai=rowSums(pij*intercept2)
  dpixi=pij*(intercept2-betai)
  mf=mf+dpixi
}
```
check with packages
```{r}
z<-with(mlogit0,data.frame(Income=tapply(Income,idx(clogit0,2),mean)))
effects(mlogit1,covariate="Income",data=z)
```
```{r}
mf/4470
```
As income goes up, choice1, chocie2, choice5, choice7 will have lower probabiltiy of getting purchased. But overall, these coeffcients are small.


## Exercise 5 Mixed Logit
```{r}
mixed_ll<-function(para){
beta1=matrix(0,nrow=1,ncol=11)
gamma1=matrix(0,nrow=1,ncol=20)
beta1[1,1]=0
beta1[1,2:10]=para[1:9]
beta1[1,11]=para[10]
gamma1[1,1]=0
gamma1[1,2:10]=para[1:9]
gamma1[1,11]=0
gamma1[1,12:20]=para[11:19]
intercept=cbind(0,matrix(rep(beta1[2:10],each=ni),ni,nj-1))/2
intercept1=cbind(0,matrix(rep(gamma1[2:10],each=ni),ni,nj-1))/2
intercept2=cbind(0,matrix(rep(gamma1[12:20],each=ni),ni,nj-1))
XB=intercept1+intercept2*cbind(all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13],all[,13])
WG=price[,2:11]*beta1[11]+intercept
e=exp(XB+WG)
se=rowSums(e)
 prob=e/se
  
  #Compute the neg log likelihood for each choice using the choice matrix
  ll=sum(Y*log(prob))
  return(-ll)
}
```

```{r}
set.seed(0)
model3 <-  optim(runif(19,-0.1,0.1),mixed_ll,method="BFGS", hessian = FALSE)
```
```{r}
model3$par
```
consider alternative specification:
```{r}
price[,12]=all$Income
colnames(price)[12] <- "Income"
logit0=mlogit.data(price,varying=2:11,shape="wide",sep="",choice="choice")
logit1=mlogit(choice~price|Income,data=logit0)
summary(logit1)
```
```{r}
#removing one choice, removing choice10 data
pricer<-price %>%
  filter(price$choice!=1)
pricer<-pricer[-c(2)]
```
```{r}
all1<-all %>%
  filter(all$choice!=1)
ni=nrow(pricer)
nj=ncol(pricer[,2:10])
Y=class.ind(all1$choice)
```
```{r}
#computing ll
mixed_ll2<-function(para){
  beta1=matrix(0,nrow=1,ncol=10)
gamma1=matrix(0,nrow=1,ncol=18)
beta1[1,1]=0
beta1[1,2:9]=para[1:8]
beta1[1,10]=para[9]
gamma1[1,1]=0
gamma1[1,2:9]=para[1:8]
gamma1[1,10]=0
gamma1[1,11:18]=para[10:17]
intercept=cbind(0,matrix(rep(beta1[2:9],each=ni),ni,nj-1))/2
intercept1=cbind(0,matrix(rep(gamma1[2:9],each=ni),ni,nj-1))/2
intercept2=cbind(0,matrix(rep(gamma1[11:18],each=ni),ni,nj-1))
XB=intercept1+intercept2*cbind(all1[,13],all1[,13],all1[,13],all1[,13],all1[,13],all1[,13],all1[,13],all1[,13],all1[,13])
WG=all1[,4:12]*para[9]+intercept
e=exp(XB+WG)
se=rowSums(e)
prob=e/se
ll=sum(Y*log(prob))
return(-ll)
}
```
```{r}
set.seed(0)
model4 <-optim(runif(17,-0.1,0.1),mixed_ll2,method="BFGS", hessian = FALSE)
```
```{r}
model4$par
```

```{r}
logit01=mlogit.data(pricer,varying=2:10,shape="wide",sep="",choice="choice")
logit11=mlogit(choice~price|Income,data=logit01)
summary(logit11)
```
```{r}
ni=nrow(price)
nj=ncol(price[,2:11])
Y=matrix(0,ni,nj)
for (i in 1:nj){
  for (j in 2:ni){
    if (price$choice[j]==i){
      Y[j,i]=1
    }
  }
}
mixed_ll(model3$par)
```
We can see that the diff between log likelihood reaches 3400, which shows that IIA has been violated.